TEMPERATURES AND METALLICITIES FOR 10,000 M DWARFS\\ IN THE \apogee\ SURVEY

### 2019-07-19 paper/draft.tex
remove figs
mark places needing revision

\begin{abstract}
We present a catalog of spectroscopic temperatures, metallicities,
spectral types for 10,311 M dwarfs in the \apogee\ survey using
\thecannon, which is a flexible, data-driven spectral-modeling and
parameter-inference framework demonstrated to estimate
stellar-parameter labels ($\teff$, $\logg$, $\feh$, and detailed
abundances) to high precision.
Using a training sample of 87 M dwarfs in the temperature range $2860
< \teff < 4130$\,K, and metallicty range $-0.5 < \feh < 0.5$\,dex
[WITH LABELS FROM WHERE??], we train a two-parameter model with
predictive accuracy (in cross-validation) to 77\,K and 0.09\,dex
respectively.
We also train a one-dimensional spectral type model using 51 M dwarfs
with known [FROM WHERE??] spectral types ranging M0 to M9, to predictive accuracy of
0.9 types.
From the overlap of \textsl{APOGEE DR14} and \textsl{Gaia DR2} we
compare our \cannon\ derived $\teff$ against color--temperature
relations and \cannon\ derived $\feh$ against isochrone models in
\gaia\ color-magnitude space.
Finally we use our results to
estimate stellar radii from absolute magnitudes and \cannon\ temperatures [AND WE FIND WHAT??].
This project has implications for the NASA \tess\ Mission and other
surveys that depend on having good parameters for M dwarfs.
\end{abstract}

@@

This paper is organized as follows: in Section \ref{sec:data} we describe the technical specifications of the data from the \apogee\ and \gaia\ surveys, as well as previous studies of M dwarfs in \apogee.
Section \ref{sec:methods} describes our model implementation using \thecannon\ framework, and Section \ref{sec:sample_selection} describes the our sample selection and derivation of training parameters.
In Section \ref{sec:results} we present our experimental results, evaluate the predictive accuracy of our models, apply our model to a selected test sample of $>$10,000 sources, and examine the validity of our parameters against color-temperature relations and isochrone grids.
Finally, in Section \ref{sec:discussion} we discuss model performance, future improvements, and implications of our results in the context of present and upcoming stellar and exoplanetary surveys. \\

@@

% ----------------------------------------
\subsection{\gaia\ Survey}

We cross-matched the {\sl APOGEE-DR14} catalog of  \citep{Brown:2018} \\

@@

The spectra contain roughly 8000 pixels,
so we might expect the $\chi^2$ values to be close to 8000 in magnitude.
They are much higher.
This discrepancy flows from the fact that---while the spectral model is good at the percent
level---the signal-to-noise of the typical spectra are more than 100.
That is, the $\chi^2$ values do show that the model is not good in the frequentist
sense; it is only good at the percent level.
This percent-level accuracy of the model in the data domain is also demonstrated... \\

@@

\begin{figure*}
	\includegraphics[width=\linewidth]{figures/aitoff.png}
	\caption{Aitoff projection of 14,827 sources.}
	\label{fig:aitoff}
\end{figure*}

@@

\begin{figure}
	\includegraphics[width=\linewidth]{figures/cannon_vs_parsec.png}
	\includegraphics[width=\linewidth]{figures/aspcap_vs_parsec1.png}
	\caption{Comparison of \cannon\ $\feh$ as compared to metallicity predicted by PARSEC isochrone tracks \citep{Bressan:2012} in \gaia\ color-magnitude space. The two panels show five metallicity tracks (ranging $-0.5<$[M/H]$<+0.5$ in steps of 0.1 dex) at two different ages (3 and 7 Gyr). In the ASPCAP plot only the 8,960 out of 10,311 sources with metallcities $-0.5<$[M/H]$<+0.5$ are shown, while the remaining 1,351 sources with [M/H]$<-0.5$ and [M/H]$>+0.5$ are shown separately in Figure \ref{fig:isochrone_cmd_2}.}
	\label{fig:isochrone_cmd}
\end{figure}

\begin{figure}
	\includegraphics[width=\linewidth]{figures/aspcap_vs_parsec2.png}
	\caption{}
	\label{fig:isochrone_cmd_2}
\end{figure}

>>>>

\begin{abstract}
We present a catalog of spectroscopic temperatures, metallicities,
spectral types for 10,311 M dwarfs in the \apogee\ survey using
\thecannon, which is a flexible, data-driven spectral-modeling and
parameter-inference framework demonstrated to estimate
stellar-parameter labels ($\teff$, $\logg$, $\feh$, and detailed
abundances) to high precision.
Using a training sample of 87 M dwarfs with optically derived labels spanning 
$2860 < \teff < 4130$\,K callibrated with bolometric temperatures, 
$-0.5 < \feh < 0.5$\,dex callibrated with FGK wide binary metallicities, 
we train a two-parameter model with predictive accuracy (in cross-validation) 
to 77\,K and 0.09\,dex respectively.
We also train a one-dimensional spectral type model using 51 M dwarfs
with known SDSS optical spectral types ranging M0 to M9, to predictive accuracy of
0.9 types.
\color{gcolor}{From the overlap of \textsl{APOGEE DR14} and \textsl{Gaia DR2} we
compare our \cannon\ derived $\teff$ against color--temperature
relations and \cannon\ derived $\feh$ against isochrone models in
\gaia\ color-magnitude space.
Finally we use our results to
estimate stellar radii from absolute magnitudes and \cannon\ temperatures [AND WE FIND WHAT??].
This project has implications for the NASA \tess\ Mission and other
surveys that depend on having good parameters for M dwarfs.}\color{black}
\end{abstract}

@@

This paper is organized as follows: in Section \ref{sec:data} we describe the technical specifications of the data from the \apogee\ and \gaia\ surveys, as well as previous studies of M dwarfs in \apogee.
Section \ref{sec:methods} describes our model implementation using \thecannon\ framework, and Section \ref{sec:sample_selection} describes the our sample selection and derivation of training parameters.
\color{gcolor}{{In Section \ref{sec:results} we present our experimental results, evaluate the predictive accuracy of our models, apply our model to a selected test sample of $>$10,000 sources, and examine the validity of our parameters against color-temperature relations and isochrone grids.
Finally, in Section \ref{sec:discussion} we discuss model performance, future improvements, and implications of our results in the context of present and upcoming stellar and exoplanetary surveys.}\color{black} \\

@@

@@

The spectra contain roughly 8000 pixels,
so we might expect the $\chi^2$ values to be close to 8000 in magnitude.
They are much higher.
This discrepancy flows from the fact that---while the spectral model is good at the percent
level---the signal-to-noise of the typical spectra are more than 100.
That is, the $\chi^2$ values do show that the model is not good in the frequentist
sense; it is only good at the percent level.
\color{gcolor}{This percent-level accuracy of the model in the data domain is also demonstrated...}\color{black} \\

@@

@@

\begin{figure}
	\includegraphics[width=\linewidth]{figures/cannon_vs_parsec.png}
	\includegraphics[width=\linewidth]{figures/aspcap_vs_parsec1.png}
	\caption{Comparison of \cannon\ $\feh$ as compared to metallicity predicted by PARSEC isochrone tracks \citep{Bressan:2012} in \gaia\ color-magnitude space. The two panels show five metallicity tracks (ranging $-0.5<$[M/H]$<+0.5$ in steps of 0.1 dex) at two different ages (3 and 7 Gyr). In the ASPCAP plot only the 8,960 out of 10,311 sources with metallcities $-0.5<$[M/H]$<+0.5$ are shown, while the remaining 1,351 sources with [M/H]$<-0.5$ and [M/H]$>+0.5$.}
	\label{fig:isochrone_cmd}
\end{figure}

% \begin{figure}
% 	\includegraphics[width=\linewidth]{figures/aspcap_vs_parsec2.png}
% 	\caption{}
% 	\label{fig:isochrone_cmd_2}
% \end{figure}
